Bayesian phylogenetic estimation of fossil ages

### 2015-09-08 fossilDating.tex
small changes
many small edits

% 250 words
\abstract{
Recent advances have allowed for both morphological fossil evidence and molecular sequences to be integrated into a single combined inference of divergence dates under the rule of Bayesian probability. 
In particular the fossilized birth-death tree prior and the Lewis-MK model of evolution of discrete morphological change allow for the estimation of both divergence times and phylogenetic relationships between fossil and extant taxa. 
We exploit this statistical framework to investigate the internal consistency of these models by estimating the phylogenetic age of each fossil in turn, within two rich and well-characterized data sets of fossil and extant species.
We find that we can accurately estimate the age of individual fossils based only on phylogenetic evidence. 
In fact in the two data sets we analyze the {\em phylogenetic age} of a fossil species is on average only $\sim 2$ My from the midpoint age of the geological strata from which it was excavated. 
The high level of internal consistency found in our analyses provides strong evidence that the Bayesian statistical model employed is a good fit for both the geological and morphological data, and provides striking evidence from real data that the framework used can accurately model the evolution of discrete morphological traits coded from fossil and extant taxa. 
We anticipate that this approach will have diverse applications beyond divergence time dating, including dating fossils that are temporally unconstrained, testing the ``morphological clock'', and for evaluating model assumptions when controversial phylogenetic hypotheses are obtained based on combined divergence dating analyses.
}

@@

Traditionally the practice has been to use one or more fossils as ``node calibrations'' by associated their geologically-derived age to a particular divergence in a molecular phylogeny. The age of the fossil is determined either by radiometric aging of strata above and/or below the strata, or more commonly by biostratigraphy. 
The difficulty lies in determining the appropriate ancestral species in the molecular phylogeny to associate the fossil with and the details of how this should be achieved within a full statistical inference framework \autocite{Ho2009, Heled2012, HeledDrummond2015}.
Once achieved, node calibration confers age estimates to the remaining ancestral divergences in the phylogenetic tree by the assumption of a strict or relaxed molecular clock \autocite{Thorne1998,thorne2005,yang2006,Drummond2006,drummond2010}.

@@

We extend this approach into the realm of discrete morphological evolution by presenting a statistical model of evolution that generates an expectation on the distribution of fossils and their morphological characters. 
This model has been previously presented in the context of divergence time dating \autocite{gavr2014,gavryushkina2015bayesian}. 
It is distinct from alternative divergence time dating approaches in that it provides an explicit treatment of the temporal information contained in fossil remains, whether or not related molecular sequence data is available. 
This leads to an estimate of the age of the most recent common ancestor of a group of fossils. 
We exploit this framework to attempt the estimation of the phylogenetic age of individual fossils based solely on morphological data and their phylogenetic affinities to related fossils of known age. 
The method is applied to two rich and well-characterized morphological data sets: (i) extant penguins and 36 fossil relatives \autocite{ksepka2010,ksepka2012}, (ii) a sample of nine extant canids and 123 fossil relatives \autocite{Slater2015}.

@@

\textcite{gavryushkina2015bayesian} described a ``total-evidence'' approach \autocite[implemented in BEAST2;][]{Beast2} for phylogenetic estimation of time-trees that employs both fossil and molecular sequence data as equal partners under the rule of probability. 
We extend their work further by investigating the consistency between the estimated phylogenetic age of a fossil and the corresponding palaeontological age range based on radiometrically-dated geological strata. 
The model of time-tree phylogeny employed is the so-called fossilized birth-death process \autocite{Heath2014}, which forms a prior probability distribution on the space of sampled-ancestor trees \autocite{Gavr2013}.

A previous study described how a set of fossils with discrete morphological characters could be used to estimate a time-tree or chrono-phylogeny. 
Here we additionally allow for one or more of the fossils to have broad uninformative priors on their age. 
This allows for the age of some of the fossils to be estimated solely based on their morphological characters and the phylogenetic affinities of their morphology to other fossils with known ages in the time-tree. 
We refer to this estimate of a fossil's age as its {\em phylogenetic age}. 
In estimating each of the fossils' phylogenetic ages in turn, three questions can be answered: (i) How much information about an individual fossil's age is available from phylogenetic analysis of morphological data alone, (ii) What is the level of phylogenetic evidence in support of the palaeontological age range for a fossil, and (iii) How does the amount of morphological data available for a new fossil and the number of related reference fossils of known age affect the accuracy of the phylogenetic estimate of a fossil's age.
These three questions are investigated using two morphological data sets, one of 36 fossil penguins and their extant relatives \autocite{ksepka2010,ksepka2012,gavryushkina2015bayesian}, and one of canids \autocite{Slater2015}.

\subsection*{Estimating the phylogenetic age of penguin fossils}

For each of the 36 penguin fossils in turn we performed a separate Bayesian phylogenetic analysis in which the focal fossil's palaeontological age constraints were replaced by a broad uniform prior of $(0,T_\text{max})$ Mya. 
\todo{This needs to change to the new prior!} 
For the penguins we chose $T_\text{max}=160$ as the upper limit as we expect no serious argument that any crown bird fossil would be older than 160Mya and it also corresponds to the upper limit on the origin of the whole phylogeny. 
For the choices of morphological substitution model and tree prior we followed \textcite{gavryushkina2015bayesian} and considered two of her modelling combinations. 
The first was \Mstrict{} in which a strict clock and the simplest model of morphological substitution was assumed \autocite{Lewis2001}.
We also employed \Mrelaxed{} from \textcite{gavryushkina2015bayesian} which involved a more complex model assuming a relaxed clock model \autocite{Drummond2006}, partitioning of the sites into groups of equal state count and an additional parameter for gamma-distributed rate variation across sites \autocite{yang:1994ma}.

\subsection*{Bayes factor}

@@

We simplify, using the equations for $p_i(t;d,r)$ given in \cite{Kendall1948}, to obtain,
\begin{eqnarray*}
p(t|T,d,r,s) &=& \frac{1}{1-p_0(T;d,r)}  \psi p_1(T-t|d,r) \\ & & \times \left[  \frac{1}{(1-p_0(T-t|d,r)/r)^2} - \frac{p_0(t|d,r)}{(1-p_0(t|d,r) p_0(T-t|d,r)/r)^2}  \right]\\
&=&  \psi  \left[ \frac{e^{d(T-t)}}{1-p_0(T;d,r)} - \frac{p_0(T|d,r)-p_0(T-t|d,r)}{(1-p_0(T-t|d,r))(1-p_0(t|d,r))} \right],
\end{eqnarray*}
with $p_0(t|d,r) = \frac{1-e^{-d*t}}{1/r - e^{-d*t}}$.

@@

\begin{figure}
\includegraphics[width=6in]{priors.pdf}
\caption{\label{hist8_older}Probability density for the sampling times under the fossilized birth-death process. The dot-dashed line uses priors on the parameters as in \textcite{gavryushkina2015bayesian}. The solid line uses the new prior with implicit assumptions on $T$ and $s$, the dashed line results from only assuming the implicit prior on $T$, the dotted line results from only assuming the  implicit prior on $s$. (Since the dashed, dotted and dot-dashed lines are governed by rare parameter combinations leading to huge trees and huge sample sizes, these lines are very sensitive to drawing another rare event, and thus need to be taken with some caution).}
\label{Fig:Prior}
\end{figure}

@@

@@

The second data set that we investigated was a morphological data matrix of 132 canid species \autocite[9 extant and 123 fossil;][]{Slater2015}. 
The 9 extant species represent about 25\% of the extant canid species and include representatives of four genera (6 {\em Canis}, 1 {\em Cuon}, 1 {\em Lycaon}, 1 {\em Urocyon}) and both tribes (8 Canini, 1 Vulpini). 
For this data set we did not have a geological age range for each fossil, but instead the author of the study provided a point estimate for each fossil species' geological age.

We applied the M1 model to this data set assuming a fixed geological age for each taxon. Unlike the original study \autocite{Slater2015} we did not apply any other constraints or priors on ancestral divergence times beyond the ages of the included fossils. 
As with the penguin data set we also performed an analysis for each of the 123 canid fossils in turn in which we relaxed the focal fossil's palaeontological age constraints to a broad uniform prior of $(0,T_\text{max})$ Mya.

@@

As judged by Bayes factors, none of the fossils exhibited strong evidence (i.e. $|\text{log BF}| > 3.0$) that the phylogenetic age was inconsistent with the geological age range assuming a uniform prior from $(0,160)$ Mya. 
However if we consider only the posterior probability that the fossil is in the geological age range then three of the 36 fossils has a posterior probability $< 0.05$, suggesting low posterior support for the phylogenetic age being within the palaeontological age range.
These three fossils were {\em Madrynornis mirandus}, {\em Paraptenodytes antarcticus} and {\em Sphenicus muizoni} with posterior probabilities that the phylogenetic age is in the palaeontological range of 0.01, 0.0011 and 0.0007 respectively. 
All other fossils has posterior probabilities of $> 0.05$ of their age being in the palaeontological range. 
It is worth noting that the absolute discrepancy in the ages are still quite moderate for the three fossils with low posterior probabilities, with {\em M. mirandus}: 6.3Myr vs 10Myr (phylogenetic age versus palaeontological age), {\em P. antarcticus}: 29.9 vs 22, {\em S. muizoni}: 5.1 vs 9.1. 
The small posterior probabilities are partially caused in these cases because the corresponding palaeontological age range is narrow, suggesting very precise geological knowledge of the ages of these three fossils.

@@

\Mrelaxed{} was the best-fitting model for this data set according to the analysis of \textcite{gavryushkina2015bayesian}. 
As with \Mstrict{} this model produced phylogenetic age estimates that were very concordant with the geological age ranges of the fossils (Figure \ref{fig:phyloAgeVsGeoAge}b), with an overall $R^2 = 0.924$. The median error was 2.1 Myr across all 36 fossils. 
Again, as with \Mstrict{}, none of the fossils exhibited strong evidence (i.e. $|\text{log BF}| > 3.0$) that the phylogenetic age was inconsistent with the geological age range assuming a uniform prior from $(0,160)$ Mya. 
However if we consider the posterior probability that the fossil is in the geological age range then five of the 36 fossils had a posterior probability $< 0.05$ for \Mrelaxed{}, suggesting low posterior support for the phylogenetic age being within the palaeontological age range. 
These five fossils were {\em Madrynornis mirandus}, {\em Paraptenodytes antarcticus}, {\em Perudyptes devriesi}, {\em Sphenicus muizoni} and {\em Waimanu manneringi} with posterior probabilities that the phylogenetic age is in the palaeontological range of 0.024, 0.016, 0.048, 0.007, 0.037 respectively. 
All other fossils has posterior probabilities of $> 0.05$ of their age being in the palaeontological range. 
Again the absolute discrepancy in the ages are quite moderate for the five fossils with low posterior probabilities, with {\em M. mirandus}: 6.5Myr vs 10Myr (phylogenetic age versus palaeontological age), {\em P. antarcticus}: 28.1 vs 22, {\em P. devriesi}: 49.0 vs 40, {\em S. muizoni}: 5.1 vs 9.1 and {\em W. manneringi}: 56.8 vs 61.05. 
A summary of all the individual estimates are tabulated in Table \ref{fossilTable8}. 
The individual marginal posterior distributions of phylogenetic age under \Mrelaxed{} and the corresponding geological range are shown in Figures \ref{hist8_younger} and \ref{hist8_older}.

@@

The canid data set shows remarkably good consistency between palaeontological ages and phylogenetic ages, even with the simple strict morphological clock model (\Mstrict{}).
The $R^2 = 0.901$ between the phylogenetic and palaeontological ages (see Figure \ref{fig:canidMorphVsGeo}). 
Only 16 out of 123 fossils (13\%) don’t have the paleontological age in the credible interval of the phylogenetic age and there are no extreme outliers. 
The median error is $1.66$My, which is even more accurate than the penguin data set.
There are less characters than the penguins (122 versus $\sim 250$) so the uncertainty in the individual age estimates is greater (median 99\% HPD range = 6.5 My), but overall it seems to work very nicely.

@@

In this paper we have demonstrated that even a small number of morphological characters (some of the fossils had as few as 7 morphological traits coded) can be used in the context of a rich fossil reference data set, to provide an accurate and precise age of the fossil based on a phylogenetic model. 
We found that although a strict morphological clock does a surprisingly good job of estimated fossil ages, there is evidence that phylogenetic estimation of fossil ages is improved by a model that includes a relaxed morphological clock.
There are a diverse array of potential applications for this methodology. 
The most obvious is the estimation of dates for fossils that are temporally unconstrained, either due to poor knowledge of the age of the sediments in which it was found or a complete lack of provenance data (e.g. a recent fossil described as a `four-legged snake' has excited controversy for a lack of provenance\footnote{See \url{http://news.sciencemag.org/paleontology/2015/07/four-legged-snake-fossil-stuns-scientists-and-ignites-controversy}}). 
It can also be used a way of testing the ``morphological clock'' and to discover potential problems in the data by identifying outlier fossils with respect to model fit. 
The median error in age estimates for the two data sets investigated were 2 My and 1.6 My respectively, using either a very simple or more complex models of discrete morphological change. 
In all cases we used the new fossilized birth-death tree prior, which is a crucial ingredient in allowing for the estimating of fossil ages under a birth-death tree prior.

\todo{How does the number of fossils in the reference affect the precision and accuracy of the fossil age estimates ?}

>>>>

% 250 words
\abstract{
Recent advances have allowed for both morphological fossil evidence and molecular sequences to be integrated into a single combined inference of divergence dates under the rule of Bayesian probability. 
In particular the fossilized birth-death tree prior and the Lewis-MK model of evolution of discrete morphological change allow for the estimation of both divergence times and phylogenetic relationships between fossil and extant taxa. 
We exploit this statistical framework to investigate the internal consistency of these models by estimating the phylogenetic age of each fossil in turn, within two rich and well-characterized data sets of fossil and extant species.
We find that we can accurately estimate the age of individual fossils based only on phylogenetic evidence. 
In fact in the two data sets we analyze the {\em phylogenetic age} of a fossil species is on average  $< 2$ My from the midpoint age of the geological strata from which it was excavated. 
The high level of internal consistency found in our analyses provides strong evidence that the Bayesian statistical model employed is a good fit for both the geological and morphological data, and provides striking evidence from real data that the framework used can accurately model the evolution of discrete morphological traits coded from fossil and extant taxa. 
We anticipate that this approach will have diverse applications beyond divergence time dating, including dating fossils that are temporally unconstrained, testing the ``morphological clock'', and for evaluating model assumptions when controversial phylogenetic hypotheses are obtained based on combined divergence dating analyses.
}

@@

Traditionally the practice has been to use one or more fossils as ``node calibrations'' by associated their geologically-derived age to a particular divergence in a molecular phylogeny. The age of the fossil is determined either by radiometric aging of strata above and/or below the fossil, or more commonly by biostratigraphy. 
The difficulty lies in determining the appropriate ancestral divergence in the molecular phylogeny to associate the fossil with and the details of how this should be achieved within a full statistical inference framework \autocite{Ho2009, Heled2012, HeledDrummond2015}.
Once achieved, node calibration confers age estimates to the remaining ancestral divergences in the phylogenetic tree by the assumption of a strict or relaxed molecular clock \autocite{Thorne1998,thorne2005,yang2006,Drummond2006,drummond2010}.

@@

We extend this approach into the realm of discrete morphological evolution by presenting a statistical model of evolution that generates an expectation on the distribution of fossils, their morphological characters. 
This model has been previously presented in the context of divergence time dating \autocite{gavr2014,gavryushkina2015bayesian}. 
It is distinct from alternative divergence time dating approaches in that it provides an explicit treatment of the temporal information contained in fossil remains, whether or not related molecular sequence data is available. 
This leads to an estimate of the age of the most recent common ancestor of a group of fossil and extant taxa. 
We exploit this framework to attempt the estimation of the phylogenetic age of individual fossils based solely on morphological data and their phylogenetic affinities to related taxa of known age. 
The method is applied to two rich and well-characterized morphological data sets: (i) extant penguins and 36 fossil relatives \autocite{ksepka2010,ksepka2012}, (ii) a sample of nine extant canids and 123 fossil relatives \autocite{Slater2015}.

@@

\textcite{gavryushkina2015bayesian} described a ``total-evidence'' approach \autocite[implemented in BEAST2;][]{Beast2} for phylogenetic estimation of time-trees that employs both fossil and molecular sequence data as equal partners under the rule of probability. 
We extend their work further by investigating the consistency between the estimated phylogenetic age of a fossil and the corresponding fossil age range determined by geological and biostratigraphic evidence. 
The model of time-tree phylogeny employed is the so-called fossilized birth-death process \autocite{Heath2014}, which forms a prior probability distribution on the space of sampled-ancestor trees \autocite{Gavr2013}.

A previous study described how a set of fossils with discrete morphological characters could be used to estimate a time-tree. 
Here we additionally allow for one or more of the fossils to have broad uninformative priors on their age. 
This allows for the age of some of the fossils to be estimated solely based on their morphological characters and the phylogenetic affinities of their morphology to other fossils with known ages in the time-tree. 
We refer to this estimate of a fossil's age as its {\em phylogenetic age}. 
In estimating each of the fossils' phylogenetic ages in turn, two questions can be answered: (i) How much information about an individual fossil's age is available from phylogenetic analysis of morphological data alone, (ii) What is the level of phylogenetic evidence in support of the palaeontological age range for a fossil.
These two questions are investigated using two morphological data sets, one of 36 fossil penguins and their extant relatives \autocite{ksepka2010,ksepka2012,gavryushkina2015bayesian}, and one of canids \autocite{Slater2015}.

\subsection*{Computing the phylogenetic evidence for an age range}

@@

We simplify, using the equations for $p_i(t;d,r)$ given in \autocite{kendall1948}, to obtain,
\begin{eqnarray*}
p(t|T,d,r,s) &=& \frac{1}{1-p_0(T;d,r)}  \psi p_1(T-t|d,r) \\ & & \times \left[  \frac{1}{(1-p_0(T-t|d,r)/r)^2} - \frac{p_0(t|d,r)}{(1-p_0(t|d,r) p_0(T-t|d,r)/r)^2}  \right]\\
&=&  \psi  \left[ \frac{e^{d(T-t)}}{1-p_0(T;d,r)} - \frac{p_0(T|d,r)-p_0(T-t|d,r)}{(1-p_0(T-t|d,r))(1-p_0(t|d,r))} \right],
\end{eqnarray*}
with $p_0(t|d,r) = \frac{1-e^{-d*t}}{1/r - e^{-d*t}}$.

@@

\begin{figure}
\includegraphics[width=12cm]{priors.pdf}
\caption{\label{hist8_older}Probability density for the sampling times under the fossilized birth-death process. The dot-dashed line uses priors on the parameters as in \textcite{gavryushkina2015bayesian}. 
The solid line uses the new prior with implicit assumptions on $T$ and $s$, the dashed line results from only assuming the implicit prior on $T$, the dotted line results from only assuming the  implicit prior on $s$. 
(Since the dashed, dotted and dot-dashed lines are governed by rare parameter combinations leading to huge trees and huge sample sizes, these lines are very sensitive to drawing another rare event, and thus need to be taken with some caution).}
\label{Fig:Prior}
\end{figure}

@@

\subsection*{Estimating the phylogenetic ages of penguin fossils}

For each of the 36 penguin fossils in turn we performed a separate Bayesian phylogenetic analysis in which the focal fossil's palaeontological age constraints were replaced by a broad model-based prior on the interval $(0,T_\text{max})$ Mya. 
For the penguin data set we chose $T_\text{max}=160$ as the upper limit as we expect no serious argument that any crown bird fossil would be older than 160Mya and it also corresponds to the upper prior limit on the origin of the whole phylogeny. 
For the choices of morphological substitution model and tree prior we followed \textcite{gavryushkina2015bayesian} and considered two of her modelling combinations. 
The first was \Mstrict{} in which a strict clock and the simplest model of morphological substitution was assumed \autocite{Lewis2001}.
We also employed \Mrelaxed{} from \textcite{gavryushkina2015bayesian} which involved a more complex model assuming a relaxed clock model \autocite{Drummond2006}, partitioning of the sites into groups of equal state count and an additional parameter for gamma-distributed rate variation across sites \autocite{yang:1994ma}.

@@

The second data set that we investigated was a morphological data matrix of 132 canid species \autocite[9 extant and 123 fossil;][]{Slater2015}. 
The 9 extant species represent about 25\% of the extant canid species and include representatives of four genera (6 {\em Canis}, 1 {\em Cuon}, 1 {\em Lycaon}, 1 {\em Urocyon}) and both tribes (8 Canini, 1 Vulpini). 
For this data set we did not have a geological age range for each fossil. Instead used the point estimates for each fossil species' geological age that were employed by the author of the original study.

We applied the M1 model to this data set assuming a fixed geological age for each taxon. Unlike the original study \autocite{Slater2015} we did not apply any other constraints or priors on ancestral divergence times beyond the ages of the included fossils. 
As with the penguin data set we also performed an analysis for each of the 123 canid fossils in turn in which we relaxed the focal fossil's palaeontological age constraints to a broad model-based prior of $(0,T_\text{max})$ Mya, where $T_\text{max} = 80$ corresponded again to the upper limit of the prior on the origin of the clade.

@@

As judged by Bayes factors, only one fossil exhibited strong evidence (i.e. $\text{log BF} < -3.0$) that the phylogenetic age was inconsistent with the geological age range. The log BF for {\em Paraptenodytes antarcticus} was -3.4. In fact the majority of the fossils (23/36 = 64\%) had strong positive evidence for the geological age range (i.e. $\text{log BF} > 3.0$).
Likewise, if we consider only the posterior probability that the fossil is in the geological age range then three of the 36 fossils has a posterior probability $< 0.05$, suggesting low posterior support for the phylogenetic age being within the palaeontological age range.
These three fossils were {\em Madrynornis mirandus}, {\em Paraptenodytes antarcticus} and {\em Sphenicus muizoni} with posterior probabilities that the phylogenetic age is in the palaeontological range of 0.007, 0.001 and 0.001 respectively. 
All other fossils has posterior probabilities of $> 0.05$ of their age being in the palaeontological range. 
It is worth noting that the absolute discrepancy in the ages are still quite moderate for the three fossils with low posterior probabilities, with {\em M. mirandus}: 6.3Myr vs 10Myr (phylogenetic age versus palaeontological age), {\em P. antarcticus}: 29.9 vs 22, {\em S. muizoni}: 5.2 vs 9.1. 
The small posterior probabilities are partially caused in these cases because the corresponding palaeontological age range is narrow, apparently suggesting very precise geological knowledge of the ages of these three fossils.

@@

\Mrelaxed{} was the best-fitting model for this data set according to the analysis of \textcite{gavryushkina2015bayesian}. 
As with \Mstrict{} this model produced phylogenetic age estimates that were very concordant with the geological age ranges of the fossils (Figure \ref{fig:phyloAgeVsGeoAge}b), with an overall $R^2 = 0.924$. The median error was 2.05 Myr across all 36 fossils. 
In this analysis none of the fossils exhibited any evidence (i.e. $\text{log BF} < 0.0$) that the phylogenetic age was inconsistent with the geological age range. 
However if we consider the posterior probability that the fossil is in the geological age range then five of the 36 fossils had a posterior probability $< 0.05$ for \Mrelaxed{}, suggesting low posterior support for the phylogenetic age being within the palaeontological age range. 
These five fossils were {\em Madrynornis mirandus}, {\em Paraptenodytes antarcticus}, {\em Perudyptes devriesi}, {\em Sphenicus muizoni} and {\em Waimanu manneringi} with posterior probabilities that the phylogenetic age is in the palaeontological range of 0.035, 0.018, 0.046, 0.004, 0.037 respectively. 
All other fossils has posterior probabilities of $> 0.05$ of their age being in the palaeontological range. 
Again the absolute discrepancy in the ages are quite moderate for the five fossils with low posterior probabilities, with {\em M. mirandus}: 6.7Myr vs 10Myr (phylogenetic age versus palaeontological age), {\em P. antarcticus}: 28.0 vs 22, {\em P. devriesi}: 49.0 vs 40, {\em S. muizoni}: 5.1 vs 9.1 and {\em W. manneringi}: 56.7 vs 61.05. 
A summary of all the individual estimates are tabulated in Table \ref{fossilTable8}. 
The individual marginal posterior distributions of phylogenetic age under \Mrelaxed{} and the corresponding geological range are shown in Figures \ref{hist8_younger} and \ref{hist8_older}.

@@

The canid data set shows remarkably good consistency between palaeontological ages and phylogenetic ages, even with the simple strict morphological clock model (\Mstrict{}).
The $R^2 = 0.901$ between the phylogenetic and palaeontological ages (see Figure \ref{fig:canidMorphVsGeo}). 
Only 16 out of 123 fossils (13\%) don't have the paleontological age in the credible interval of the phylogenetic age and there are no extreme outliers. 
The median error is $1.66$My, which is even more accurate than the penguin data set.
This data set contains half as many morphological characters as does the penguin data set (122 versus $245$) but despite that the individual age estimates have more precision (median 99\% HPD range = 6.5 My for canids as opposed to 9.6My for penguins).

@@

In this paper we have demonstrated that even a small number of morphological characters (some of the fossils had as few as 7 morphological traits coded) can be used in the context of a rich fossil reference data set, to provide an accurate and precise age of the fossil based on a phylogenetic model. 
We found that although a strict morphological clock does a surprisingly good job of estimated fossil ages, there is evidence that phylogenetic estimation of fossil ages is improved by a model that includes a relaxed morphological clock. 
We found that fossil estimates were both slightly more accurate and more precise on average in the canid data set. 
This may suggest that the larger reference set of fossils in the canid data set (122 versus 35) more than makes up for the smaller number of characters (122 versus 245) with regards to accuracy and precision of fossil age estimates.

There are a diverse array of potential applications for this methodology. 
The most obvious is the estimation of dates for fossils that are temporally unconstrained, either due to poor knowledge of the age of the sediments in which it was found or a complete lack of provenance data (e.g. a recent fossil described as a `four-legged snake' has excited controversy for a lack of provenance\footnote{See \url{http://news.sciencemag.org/paleontology/2015/07/four-legged-snake-fossil-stuns-scientists-and-ignites-controversy}}). 
It can also be used a way of testing the ``morphological clock'' and to discover potential problems in the data by identifying outlier fossils with respect to model fit. 
The median error in age estimates for the two data sets investigated were 2 My and 1.6 My respectively, using either a very simple or more complex models of discrete morphological change. 
In all cases we used the new fossilized birth-death tree prior, which is a crucial ingredient in allowing for the estimating of fossil ages under a birth-death tree prior.

Future work may investigate in a more systematic fashion how the amount of morphological data available for a new fossil and the number of related reference fossils of known age affect the accuracy of the phylogenetic estimate of a fossil's age.

In the mean time we anticipate that this approach will help to promote the application of a consistent probabilistic framework to consider both molecular and fossil evidence. Our results are encouraging in suggesting that the statistical models presented are adequate for inference of phylogenetic time-trees from morphological fossil data.
